\plaintitle

### 2016-05-23 chi/papers/2017/referrals/referrals.tex
some revisions on core argument; cleaned up referrals.tex

\begin{abstract}
  With growing attention on gig work
  --- ranging from the ``sharing economy'' to microtasks ---
  scholars have made connections to frameworks like Taylorism,
  and mechanisms such as worker advocacy and empowerment,
  to make sense of our observations of
  on--demand work and the workers that power this movement.
  We argue that our the underlying trend towards ``piecework''
  --- driven in part by the discretization,
  routenization,
  and external management of said work ---
  not only suggests, but in fact generates what we have observed:
  members of this transient workforce increasingly feeling
  disempowered,
  marginalized,
  and frustrated by the systems and platforms on which they work.

  After evaluating this framing
  through a series of case studies
  in various industries falling broadly under the ``gig work'' category,
  we turn our theoretical lens to look to the future,
  to identify worthwhile questions and
  points of inquiry that researchers in social computing should consider
  as we attempt to anticipate and perhaps shape the future of work.
\end{abstract}

@@

\section{Introduction}
This section should outline the trend of contemporary gig work in the last 10 years,
in particular tracking the changes since
\citeauthor{crowdworkFuture} discussed
\citetitle{crowdworkFuture}
in \citeyear{crowdworkFuture}
\cite{crowdworkFuture}.
This would review the literature since then,
then point to the movements toward worker unionization, advocacy, etc\dots
that have happened in the last year or two.

In the reflection on the literature published in the last 5 years since
\citetitle{crowdworkFuture},
we notice a broader change in the form of work that's being done;
motivated by this realization,
we look at the emergent phenomena in piecework both as interconnected.
By looking at the movements toward
discretization,
routenization,
and management
as necessarily linked --- even causally --- to one another, 
we give ourselves a framing of contemporary piecework that explains
and arguably predicts
what we have seen thus far.

\subsection{Piecework as a lens to understand gig work}
It's not new that researchers have discussed gig work as an extension or resurgence
of piecework, but it is sufficiently important for the framing of this argument that
it merits a more direct treatment and some explanation here.
\citeauthor{crowdworkFuture}
investigate the future of crowd work by
situating and interrogating it through ``piecework'',
a term almost lost to history, but which for a time described
work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration})
\cite{crowdworkFuture}.
This work was largely in textiles and was predominantly done by women,
which by itself is a topic worthy of unpacking
\cite{scott1975women}.
Given the scope, however,
we will have to leave this complex topic with the superficial takeaway that
piecework and microtasks share some structural similarities:
\begin{inlinelist}
\item the work is done in the home;
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
\end{inlinelist}

% The work involved often meant sewing together pieces of denim,
% was largely done by women, and turned

Crowd work,
they argued in \citeyear{crowdworkFuture},
represents a threat to the future of work inasmuch as it marginalized
and perhaps even harmed
workers;
their driving question, then, was
whether the future of
(crowd, but arguably all discretized)
work might be one in which they wish their children to participate someday;
it seems strikingly like one that might have been asked
as patches of denim were first being delivered to the homes of early pieceworkers.

This paper will argue that the similarity is more than striking;
it was an inevitable milestone in the maturation of the type of work that is reemerging today.
\textit{How to continue???}

% we engaged in ethnographic fieldwork
% to attempt to understand why gig work appeals to those that engage in it,
% and to gain a better understanding of how they relate to the marketplaces in which they work.

% \textit{I'd also like to bring up piecework here,
% and briefly make the case that
% the kind of work we're seeing should be thought of as a continuation
% or resurgence
% of piecework, with some important distinctions that differentiate it.}

\subsection{Context}
Importantly, since the reemergence of piecework we've seen substantial frustration
and resistance among the workers in this area
\cite{uberAlgorithm,turkopticon,dynamo}.
This paper attempts to make sense of the broader research on this piecework,
or ``gig work'', by framing this as one of several steps in the marginalization of workers,
starting with the discretization of tasks,
followed by routenization and the rise of workflows,
and finally the external management of workers.
All of this is to say that these milestones follow sequentially,
not coincidentally but necessarily,
and that by tracing this path using the corpus of scholarship on labor and workers
we can both make sense of past events and perhaps reasonably predict next steps.

We explore each topic
--- discretization, routenization, \& management ---
by looking at case studies in social computing.
Having validated this lens as a way of reasoning about contemporary piecework, or ``gig work'',
we turn to look ahead, envisioning future areas that researchers in social computing
--- and particularly digitally mediated work ---
should explore.

\section{Case Studies}
The past decade of research has shed light on crowd work from various perspectives,
and revealed 
\subsection{Discretization}
\textit{I want to argue that modern gig work is discretized,
and that this parallels some of the piecework that we're historically familiar with}
Work has become discretized in the last 10 years, particularly in the information work sector.
Breakthroughs like ``Find--fix--verify'', ``Cascade'', etc\dots
make it easier to assign smaller components of work to myriad workers and
recompose the constituent parts into something more complex
\cite{bernstein2015soylent,chilton2013cascade}.
Some work has even turned to ``break it down'',
inasmuch as decomposing what we might call ``macro'' tasks into microtasks
\cite{cheng2015break}
(\textit{there's a lot more to talk critique from this paper; sorry Justin.}).

The parallels between this kind of work and
historical piecework of the turn of the \nth{20} century
is by no means new
[\textit{citation needed --- who said this first? most notably? Turner?}].
\textit{Elaborate on gig work, piecework, etc\dots}

\subsection{Routenization}
The \textbf{discretization} and decomposition of work certainly facilitates
breaking tasks down,
parallelizing work,
and getting more done across a broader array of people;
but the key advantage in turning macro tasks 
into many micro tasks is that we can make that work \textit{routine}.

By making larger bodies of work less contextually situated,
we can define processes that make that work
and the instructions thereof
useful for virtually anyone with shared cultural or intellectual background.

Routenization can be seen in the contemporary in workflows
\cite{foundry,bernstein2015soylent}
[\textit{There are various papers that talk about creating a workflow;
maybe this is where FFV belongs? Foundry?}].

\textit{Harken back to piecework
and the instructions that denim workers had to make clothing;
the instructions were clear and scoped enough that you could repeat the task ad infinitum.}

\subsection{External Management}
\textbf{Discretization}, enabling \textbf{routenization}, thus allows us to
evaluate work in abstraction.
This means that we can

\textit{I want to cite Justin's work on crowdsourcing effort to highlight that
we can evaluate workers for the work they do with increasingly finite amounts of time;
also, Ranjay's work that embraces failure because
we can essentially algorithmically make sense of failure (let's call it noise)
and re--capture signal.}

\subsection{Resistance}
So it's not surprising, then, that workers are frustrated:
Turkopticon, Dynamo, and many other pieces of work point this out
\cite{turkopticon,dynamo,uberAlgorithm}
[\textit{also want to cite Brian's work from CHI this year}].
In popular culture, reporters have inquired about
what ``the future of labor unions'' will look like
[\textit{Imagine I had cited an online article and
it wasn't totally inappropriate for a conference}].

\textit{The parallels here would be difficult to ignore;
Lily and others have pointed this out,
but it's worth thinking about the broader trend of
resisting management,
driven by routenization,
enabled by discretization.
Do that here.}

\subsection{Looking forward}
If we agree that
discretization,
routenization,
management,
and even resistance
necessarily follow one another according to this theoretical lens,
then we have to use it to attempt to envision
what comes next.

\textit{I've been going back and forth regarding how to frame this paper;
the approach that tries to look at
things that are similar vs things that are different
doesn't seem to work, but
I've left the thoughts here because it's not all completely bad.}

\section{Things stay the same}
How is gig work the same as it's been historically?

I think this section would be compelling to draw parallels between
the narratives drivers gave about the flexibility, autonomy, etc\dots
and that which we might have seen among pieceworkers
(predominantly women, who benefited from being able to work from home).

\subsection{Flexibility}
Are there cultural differences between the people that did piecework and
the people that do gig work now?
I'm not sure there are significant differences that have affected the outcomes so far.

Many of the workers to whom piecework appealed were mothers, wives, etc\dots
who mostly stayed at home for various reasons
(certainly largely it was cultural
--- women weren't afforded equal access to labor opportunities,
making in--home job opportunities not only compelling, but also one of few available options).

Gig workers are in some senses similarly constrained:
workers on Amazon Mechanical Turk
--- those that use it as a primary source of income, at least ---
report being homebound for various reasons (e.g. medical, parenting, etc\dots).
Society and circumstance have made it difficult
or impossible
to join the contemporary, conventional workforce;
gig work re--opens that door.

But there are differences; during our research over the summer,
we spoke to drivers on Uber and Lyft, cleaners, and other gig workers.
Many of them told us about their home lives
--- about children, spouses, and other commitments ---
to which they wanted to dedicate more time.
One driver
(let's call him Ra\'{u}l)
told me about how he drove for Lyft
after working as an inventory manager at a hospital for more than nine years.

I asked him why he quit that job and forewent
the benefits,
predictability,
and career growth opportunity
that his old job offered.
He told me that when his daughter was born,
he was overcome with a desire to spend more time with her.
No longer satisfied with work
where he often left before his daughter would wake up and return after she fell asleep,
Ra\'{u}l decided to start driving for Lyft,
because he could drive in the evenings when his daughter was asleep.

Other drivers reported similar benefits;
gig work affords its workers flexibility that conventional careers don't allow.

\subsection{Routenization of work}
This leads into the next section, but I want to bring up the process of
making work about mass--manufacturing, at least inasmuch as
the instructions are the same for everyone.

The Internet arguably has made it much easier to broadcast those instructions,
but it's had this deeper effect of enabling some amount of back--and--forth
between the worker and the (algorithmic) manager.

\subsection{Taylorism}
The routenization of work makes it possible to measure that process,
optimize it for certain characteristics, and ultimately
lead to Taylorism and scientific management.
This is not new;
researchers have studied and written about the slow creep of
algorithmic management and 
discretization \& routenization of work tasks.

\textit{We hope to take a step back from the context in which this work is often applied, and
look for its place in the larger trends and theories
to make sense of the trends of gig work at large.}

Industrialization and the automobile assembly line makes this famous,
but piecework functioned on the principle that everyone was making similar
or identical
garments and other products.

Now, we see Turkers being evaluated on the outcome of their work conforming to norms,
sometimes bootstrapped,
as in Ranjay's talk on 
``Embracing Error to Enable Rapid Crowdsourcing'',
but more conventionally in work flows like ``Find--Fix--Verify''.

\textit{This might be an opportunity to reflect on
how pieceworkers internalized the work they were doing,
responded to the stress of the uncertainty of potentially rejected work,
etc\dots but I'm not familiar with research in that space.}

\subsection{The emergence of decentralized workplaces}
The practice of in--home piecework was consumed by
the centralization of factories
(the effect of which we'll talk about later, since
we can talk about how this made unions more practical),
but for a time many pieceworkers
at the turn of the \nth{20} century
worked out of their homes.
Strikingly, many of the gig workers
\citeauthor{crowdworkFuture}
discussed in
\citeyear{crowdworkFuture},
and indeed many more continue to do what we might call ``information work''
--- that is, work that predominantly demands human computation ---
but increasingly we're seeing the movement toward transient work that largely requires embodied presence
\cite{uberAlgorithm} (and others).

\subsection{What was this about?}
I was going to check in quickly and take a photo of your whiteboard
(hoping that you'd kept our conversation notes around)
but your room was \textit{literally} full of people lol.

\section{Things are changing}
The medium on which this work is being done
--- and to an extent the medium used to manage workers ---
has dramatically changed things as well, however;
workers are distributed around the world,
working out of their cars in the cases of livery services
(notably, never returning to a base of operations)
across and between cities as well as nations,
or in their homes
(paralleling the trend of piecework even more closely).

Trying to understand how gig work has differed from piecework should at least
start with looking at the different characteristics of the work involved.
After that, we should think about how
the demography and culture of the people engaging in this kind of work
have changed versus that of the pieceworkers.

\subsection{Differences in the work itself}
Gig work has all of the above similarities with piecework,
but there are key differences.

Piecework emerged at the turn of the \nth{20} century,
right at the time that telecommunications began to boom in the United States. % too far a reach?
\textit{Did telecom enable remote management?}

Gig work in its contemporary formation is largely mediated by
ubiquitously accessible digital media
(the Internet, telephony, etc\dots)
and importantly has relied on this technology to facilitate the remote management of workers
\cite{uberAlgorithm}.

\textit{Has the work fundamentally changed,
or are we just being managed remotely in different ways?
I'm not entirely sure.}

How does this work differ from the experience of being ``on--call'' that is so familiar to retail employees?
How does this work differ from the work in which truckers, taxi drivers, and other
independent contractors have been participating for decades?
We argue that the substantive difference in these markets is the speed of the market itself,
motivated by the technology which mediates it.
Because workers can be sourced and dispatched virtually instantly,
businesses that engage in this kind of work
(e.g. Uber, Amazon Mechanical Turk, etc\dots)
have taken to removing other bottlenecks,
like vetting workers upfront

\section{More deliberate work}
% \subsection{Differences in the people that do the work}
% ???

>>>>

\begin{abstract}
  This paper is going to be about giving workers an opportunity to refer other workers.
  Some of the old literature on job referring suggests that
  it's a good way for networked workers to do stuff.
  \citeauthor{crowdcollab} found that \citetitle{crowdcollab},
  so why not take advantage of that?

  We'll try various incentive structures
  (no penalty/no rejection, task approval, referral commission, and referral bonus)
\end{abstract}

@@

Hi there.
\section{Introduction}
The practice of referring work to other skilled workers has been well--studied in economics and more broadly in the social sciences for decades
[\textit{citation needed}].
But researchers in crowd work haven't sufficiently explored the potential
to employ this characteristic of socially connected workers
to give professionals opportunities to better route tasks.
One possible reason for this may have been the assumption that workers are not necessarily well--connected.
This would be an intuitively reasonable assumption
--- with no in--built forums or communities for workers on various gig labor markets
(e.g. Uber, Amazon Mechanical Turk (AMT), etc\dots),
one might reasonably conclude that these workers are not collaborative.

But, as
\citeauthor{crowdcollab} identifies,
crowd workers are substantially more collaborative than we originally may have inferred
\cite{crowdcollab}.
Given this realization, we pose the question:
can requesters operationalize this finding to get work done more effectively than conventional methods would allow?
Perhaps more importantly, can employing this finding allow workers
to contribute meaningfully to tasks
in ways that do not necessarily fall under the existing paradigms
(e.g. ``find--fix--verify'')
\cite{bernstein2015soylent}?

### 2016-05-23 cscw/papers/2017/labor/trackingWork.tex
some revisions on core argument; cleaned up referrals.tex

% End of preamble. Here it comes the document.
\begin{document}
\balance{}
\title{\plaintitle}

@@

The past decade has seen
microtasks, the ``sharing economy'', and other instantiations of on--demand contract work
grow to occupy the fascination of both academic circles and our culture as a whole
\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
The research community has identified myriad connections between this emergent form of work
and the historically situated ``piecework'',
\msb{Did the research community really do this? Or is \textbf{your} contribution? I want to understand whether the literature threw around the word as a metaphor, or if they really meant it in the full historical context of the term.}
as well as a number of observations relating to
the frustration workers feel stemming from the management of this work
\cite{turkopticon,dynamo,uberAlgorithm}.
This body of work represents the cornerstones of how we understand crowd workers and on--demand labor more generally.

While much of this work appears to gesture toward
the parallels between contemporary on--demand work and piecework,
it's proven difficult to bring the totality of these observations into focus using one theoretical lens.
This paper will attempt to do so by arguing that
the topics social computing researchers have investigated are
not just parallel to historical piecework and
the process of factorization that took place in the early \nth{20} century;
indeed, these phenomena are directly related \msb{how? what's the difference between parallel and directly related?}, and that they \msb{`they' means `the topics social computing researchers have investigated'? I'm not sure I understand which topics these are, or how they precipitate outcomes} precipitate
the social outcomes we have observed,
especially with regard to worker frustration, coordination, and resistance.
\msb{At this point I'm not sure what the argument is. I see several possible arguments being made: 1) People already talked about piecework, and I'm going to talk about it more and argue its `direct' relation as opposed to `parallel' relation. 2) People have talked about the \textit{effects} of gig work, and I'm going to tie them all together under one \textit{cause}---piecework---and show how that cause led to them historically and will now too. 3) There are topics that social computing researchers investigated (not sure which) and they are directly related to these outcomes. I sort of assumed it was going to be number two, but that's not what I understood here.}

% This section should outline the trend of contemporary gig work in the last 10 years,
% in particular tracking the changes since
% \citeauthor{crowdworkFuture} discussed
% \citetitle{crowdworkFuture}
% in \citeyear{crowdworkFuture}
% \cite{crowdworkFuture},.
% In particular, we point to \citeauthor{crowdworkFuture}'s discussion of worthwhile
% questions to ask regarding \citetitle{crowdworkFuture}
% \cite{crowdworkFuture}.
% This would review the literature since then,
% then point to the movements toward worker unionization, advocacy, etc\dots
% that have happened in the last year or two.

In the reflection on the literature published in the last 5 years since
\citeauthor{crowdworkFuture}'s 
\citetitle{crowdworkFuture},
we notice a broader trend describing the change in work that's being done
% motivated by this realization,
% we look at the emergent phenomena in piecework both as interconnected
\cite{crowdworkFuture}.
Crowd--powered work is not just distributed across homes,
but dispersed among drivers, cleaners and other on--demand workers \msb{I don't understand the distinction. Was Kittur et al's vision about homes somehow? Or are you saying that it was about information work but in fact is occuring elsewhere?}.
By looking at the movements toward
distributing work,
routinizing and breaking down tasks,
and externally managing workers
% discretization,
% routinization,
% and management
as linked to one another, 
we give ourselves a framing of contemporary piecework that explains,
and arguably predicts,
what we have seen thus far
--- and perhaps what we should expect to see going forward.
\msb{this paragraph sounded like half review of previous literature, and then the second half switched into ``my contribution'' mode. Either a clearer transition or a more singular purpose would help.}

% The work involved often meant sewing together pieces of denim,
% was largely done by women, and turned

Crowd work,
they argued in \citeyear{crowdworkFuture},
represents a threat to the future of work inasmuch as it marginalized
and perhaps even harmed
workers;
their driving question, then, was
whether the future of
(crowd, but arguably all discretized)
work might be one in which they wish their children to participate someday;
it seems strikingly like one that might have been asked
as patches of denim were first being delivered to the homes of early pieceworkers. \msb{this seems speculative---I'm not sure whether they knew what was coming. and wasn't income distribution different then? Who was the `they' in this case? Industrialists?}

This paper will argue that the familiarity is more than passing;
it \msb{what is `it'?} was an inevitable milestone in the maturation of the type of work that is reemerging today.
\msb{I'd like the argue to summarize the argument as to \textbf{why} it was inevitable. Why was it inevitable then, and why is it inevitable now? What is it about piecework that generates this outcome? In two sentences or less...}
``But'',
as \citeauthor{scholz2012digital} points out,
``it would be wrong to conclude that
in the realm of digital labor there is nothing new under the sun''
\cite{scholz2012digital}.
Indeed, we will attempt to reason about more than the larger--scale progression from
the discretization of work toward scientific management and future steps;
we will attempt to make sense of how the modern processes in digital labor,
in particular the use of the Internet as a medium to manage as well as \textit{do} work,
affect these unfolding events.
\msb{That felt like a bit of a calorie-free sentence --- I'm not sure what it was trying to say. I think it needs to connect the dots more visibly for the reader. Is it that we're using histories and theories of piecework to reason about that progression as it applies to today?}
\msb{I'd like to see a 1--2 sentence summary of what the takeaways from this part of the exploration are. What is new under the sun?}

% we engaged in ethnographic fieldwork
% to attempt to understand why gig work appeals to those that engage in it,
% and to gain a better understanding of how they relate to the marketplaces in which they work.

% \textit{I'd also like to bring up piecework here,
% and briefly make the case that
% the kind of work we're seeing should be thought of as a continuation
% or resurgence
% of piecework, with some important distinctions that differentiate it.}

\subsection{Piecework as a lens to understand gig work}
% It's not new to discuss gig work as an extension or even a resurgence of piecework
% \cite{crowdworkFuture,fixingChaos}.
% Nevertheless,
% it's sufficiently core to the argument here that a review is worthwhile.
\citeauthor{crowdworkFuture}
investigate the future of crowd work by
situating and interrogating it through ``piecework'',
\msb{I would not say that piecework was exactly the centerpiece of that paper. It was used as a point of contrast.}
a term almost lost to history \msb{a bit exaggerated}, but which for a time described
work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration})
\cite{crowdworkFuture} \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
This work was largely in textiles and was predominantly done by women,
which by itself is a topic worthy of unpacking \msb{instead of gesticulating, can we indicate what the punchline of unpacking it was?}
\cite{scott1975women}.
Given the scope, 
we'll leave this topic with the superficial takeaway that
piecework and microtasks share some structural similarities:
\msb{may be just me, but that felt like it was belittling the reader, like their puny brains wouldn't be able to understand more than a superficial takeaway}
\begin{inlinelist}
\item the work is done in the home \msb{true? isn't it still piecework if it's in a factory?};
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy \msb{my first reaction was ``this isn't definitional''. But now I look back and see that you're just saying there are similarities. It would be helpful if the sentence made it clearer that the similarities were in not just what the work was but also in the environment around it.}
\end{inlinelist}
It may be said that
on--demand work shares, if not inspiration \msb{I doubt it was inspiration --- `hey, let's create something like factory floors but for information work!'}, then many of the same outcomes as piecework. \msb{what are outcomes?}

@@

\textit{There's a reference to something that's \textbf{not} 99designs,
that describes a call for scientific solutions/designs to solve some problem
that the requesters have.
Honestly it's killing me that I can't remember the name,
and I need to find this before I do anything productive ever again.}
\msb{netflix prize?}

The HCI community is perhaps most familiar with examples such as
99designs,
Uber, Lyft,
and Amazon Mechanical Turk (AMT),
which all allow requesters in various forms to tap into
resources such as
cars,
computers, and above all
``cognitive surplus''
with relative ease
\cite{DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
This insight, that
workers can be geographically distributed,
% (especially given careful central management of those workers, discussed later),
has proven remarkably compelling
\cite{sensitiveTasks}.

@@

\printbibliography

>>>>

% End of preamble. Here it comes the document.
\begin{document}
\title{\plaintitle}

@@

The past decade has seen
microtasks, the ``sharing economy'', and other instantiations of on--demand contract work
grow to occupy the fascination of both academic circles and our culture as a whole
\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
The research community has made connections between this emergent form of work
and the historically situated ``piecework'',
as well as a number of observations relating to
the frustration workers feel stemming from the management of this work
\cite{turkopticon,dynamo,uberAlgorithm}.
But the connections between
historical piecework and
contemporary crowd work
haven't been deeply made as yet.
% um, note to self: say more.

While much of this work appears to gesture toward
the parallels between contemporary on--demand work and piecework,
it's proven difficult to bring the totality of these observations into focus using one theoretical lens.
This paper will attempt to do so by arguing that
the topics social computing researchers have investigated are
not just parallel to historical piecework and
the process of factorization that took place in the early \nth{20} century;
indeed, these phenomena precipitate and reinforce each other.

In the reflection on the literature published in the last 5 years since
\citeauthor{crowdworkFuture}'s 
\citetitle{crowdworkFuture},
we notice a broader trend describing the change in work that's being done
% motivated by this realization,
% we look at the emergent phenomena in piecework both as interconnected
\cite{crowdworkFuture}.
Following the improvements in
breaking down tasks,
delegating work, and
managing workers more broadly,
Crowd--powered has continued to parallel historical piecework's trajectory
by outgrowing ad--hoc worker groups and coalescing into more formal working groups;
we might call this the beginning of factorization.
% "Factorization" is not the right word but I honestly have no idea what the right term is.
% The process of formalizing work into factories? Should I tweet/post/ask about this somewhere?

% Crowd work,
% they argued in \citeyear{crowdworkFuture},
% represents a threat to the future of work inasmuch as it marginalized
% and perhaps even harmed
% workers;
% their driving question, then, was
% whether the future of
% (crowd, but arguably all discretized)
% work might be one in which they wish their children to participate someday;
% it seems strikingly like one that might have been asked
% as patches of denim were first being delivered to the homes of early pieceworkers.
% % \msb{this seems speculative---I'm not sure whether they knew what was coming. and wasn't income distribution different then? Who was the `they' in this case? Industrialists?}
% agreed; i'm commenting it out for now to see how I feel without making this argument at all. If I really feel like it's missing, I should find evidence that contemporaries discussed it...

Making sense of this 
by looking at the movements toward
distributing work,
routinizing and breaking down tasks,
and externally managing workers
as linked to one another, 
we give ourselves a framing of contemporary piecework that explains,
and arguably predicts,
what we have seen thus far
--- and perhaps what we should expect to see going forward.

This paper will attempt to demonstrate that the familiarity is more than passing;
that the phenomena researchers have observed in on--demand digitally mediated labor markets
were inevitable milestones in the birth and life of piecework.
We will set out to show that piecework itself
--- and the features inherent to it (such as decomposition of tasks) --- % and others?
precipitated the negative outcomes that historically led workers to collective organization and action in the early \nth{20} century;
that aspects of piecework like 
% Indeed, the model of work distribution and management we know as piecework
% --- including the methods and work--flows that facilitate this kind of work ---
% ultimately cause what we recognize broadly as
% frustration,
% alienation, and
% resistance
% among gig workers.

% \msb{I'd like the argue to summarize the argument as to \textbf{why} it was inevitable. Why was it inevitable then, and why is it inevitable now? What is it about piecework that generates this outcome? In two sentences or less...}
``But'',
as \citeauthor{scholz2012digital} points out,
``it would be wrong to conclude that
in the realm of digital labor there is nothing new under the sun''
\cite{scholz2012digital}.
For one, information work
(e.g. AMT \& Upwork)
and digitally managed work in general make it substantially easier
to keep workers geographically dispersed
compared to the factory workers that followed home--based piecework.
This aspect of work renders workers virtually invisible to the task solicitors
--- or ``requesters'' ---
despite being significantly more connected than historical pieceworkers were % this is a weak point...
\cite{turkopticon}.
Secondly, while
factory foremen and other middle--men
traditionally mediated the relationships between workers and managers,
today the visible agents are \textit{systems};
rather than employing individuals who ``personified the functions of management''
and can thus negotiate workers' needs,
socio--technical systems mediate these interactions % often with unrewarding outcomes
\cite{wray1949marginal}
% \cite{storiesIraniSilberman}.
Finally, this characteristic of modern piecework makes
worker coordination for collective advocacy and action significantly more difficult,
necessitating special consideration to maintain the inertia of collective will
while focusing that energy productively
\cite{dynamo}.
% here are undoubtedly more examples of differences.
% How do I not stake a claim that this is comprehensive?
% Or should I really flesh this out and make it (comprehensive)?

% Indeed, we will attempt to reason about more than the larger--scale progression from
% the discretization of work toward scientific management and future steps;
% we will attempt to make sense of how the modern processes in digital labor,
% in particular the use of the Internet as a medium to manage as well as \textit{do} work,
% affect these unfolding events.
% \msb{That felt like a bit of a calorie-free sentence --- I'm not sure what it was trying to say. I think it needs to connect the dots more visibly for the reader. Is it that we're using histories and theories of piecework to reason about that progression as it applies to today?}
% \msb{I'd like to see a 1--2 sentence summary of what the takeaways from this part of the exploration are. What is new under the sun?}

\subsection{Piecework as a lens to understand gig work}
\citeauthor{crowdworkFuture}
investigate the future of crowd work by
situating and interrogating it through ``piecework'',
\msb{I would not say that piecework was exactly the centerpiece of that paper. It was used as a point of contrast.}
a term almost lost to history \msb{a bit exaggerated}, but which for a time described
work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration})
\cite{crowdworkFuture} \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
This work was largely in textiles and was predominantly done by women,
which by itself is a topic worthy of unpacking \msb{instead of gesticulating, can we indicate what the punchline of unpacking it was?}
\cite{scott1975women}.
Given the scope, 
we'll leave this topic with the superficial takeaway that
piecework and microtasks share some structural similarities:
\msb{may be just me, but that felt like it was belittling the reader, like their puny brains wouldn't be able to understand more than a superficial takeaway}
\begin{inlinelist}
\item the work is done in the home \msb{true? isn't it still piecework if it's in a factory?};
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy \msb{my first reaction was ``this isn't definitional''. But now I look back and see that you're just saying there are similarities. It would be helpful if the sentence made it clearer that the similarities were in not just what the work was but also in the environment around it.}
\end{inlinelist}
It may be said that
on--demand work shares, if not inspiration \msb{I doubt it was inspiration --- `hey, let's create something like factory floors but for information work!'}, then many of the same outcomes as piecework. \msb{what are outcomes?}

@@

% \textit{There's a reference to something that's \textbf{not} 99designs,
% that describes a call for scientific solutions/designs to solve some problem
% that the requesters have.
% Honestly it's killing me that I can't remember the name,
% and I need to find this before I do anything productive ever again.}
% \msb{netflix prize?}
% that's an interesting example in the same vein as what i'm thinking about, but this is a whole business of it
% \textbf{INNOCENTIVE} oh my gosh it just clicked my life is complete right now at 9:54am.
% I don't even need to submit anymore.
% I'm satisfied having pulled this strand of a memory out of my brain.

The HCI community is perhaps most familiar with examples such as
Innocentive,
Uber, Lyft,
and Amazon Mechanical Turk (AMT),
which all allow requesters in various forms to tap into
resources such as
cars,
computers, and above all
``cognitive surplus''
with relative ease
\cite{howe2006rise,DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
This insight, that
workers can be geographically distributed and tasks decomposed,
% (especially given careful central management of those workers, discussed later),
has proven remarkably compelling
and an effective fulcrum for leveraging the Internet for highly scalable work
\cite{sensitiveTasks,embracingErrorKrishna}.

@@

\balance{}
\printbibliography
